% Maastricht University Preface
%
% Developed for my Master Thesis at Maastricht University.
% Based on Eugenio Senes's template at the University of Torino.
%
% By Joeri Hermans (joeri@joerihermans.com)
%
% Released under an MIT license. Share, modify and enjoy, but quote the author!

\newpage
\chapter*{Preface}
\addcontentsline{toc}{chapter}{Preface}

This thesis is submitted as a final requirement for the Master of Science degree at the Department of Data Science \& Knowledge Engineering of Maastricht University, The Netherlands. The subject of study originally started as a pilot project with Jean-Roch Vlimant, Maurizio Pierini, and Federico Presutti of the EP-UCM group (CMS experiment) at CERN. In order to handle the increased data rates of LHC Run 3 and High Luminosity LHC, the CMS experiment is considering to construct a new architecture for the High Level Trigger based on Deep Neural Networks. However, they would like to significantly decrease the training time of the models as well. This would allow them to tune the neural networks more frequently. As a result, we started to experiment with various state of the art distributed optimization algorithms. Which resulted in the achievements and insights presented in this thesis.\\

I would like to express my gratitude to several people. First and foremost, I would like to thank my promotors, Gerasimos Spanakis, and Rico M\"ockel for their expertise and suggestions during my research, which drastically improved the quality of this thesis. Furthermore, I would also like to thank my friends, colleagues and scientists at CERN for their support, feedback, and exchange of ideas during my stay there. It was a very motivating and inspiring time in my life. Especially the support and experience of my CERN supervisors, Zbigniew Baranowski, and Luca Canali, was proven to be invaluable on multiple occasions. I would also like to thank them for giving me the personal freedom to conduct my own research. Finally, I would like to thank my parents and grandparents who always supported me, and who gave me the chance to explore the world in this unique way.

\vspace{1cm}
\begin{flushright}
Joeri~R.~Hermans\\
Geneva, Switzerland 2016 - 2017
\end{flushright}
